### 2024-05-28 presentations/aps_24/aps_24.qmd
updating with APS presentation

---
title: 'Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning'
author: "Erin M. Buchanan"
institute: "Harrisburg University"
format: 
  revealjs:
    theme: night
editor: source
incremental: true 
scrollable: true 
preview-links: true
code-copy: true 
highlight-style: github 
editor_options: 
  chunk_output_type: inline
---

## Power and Sample Size Planning{.smaller}

```{css}
h1.title {
font-size: 1.5em;
}

.reveal.smaller .slides h2, .reveal .slides section.smaller h2 {
  font-size: 40px;
}

.reveal div.sourceCode pre code {
  min-height: 100%;
  font-size: 25px;
}

.reveal .cell-output-stdout pre code{
  font-size: 20px;
}
```

- Sample Size Planning: New Tools and Innovations
  - Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning, Erin M. Buchanan
  - Power Analyses for Interaction Effects in Observational Studies, David A. Baranger
  - Empowering Sample Size Justification with the Superpower R Package, Aaron Caldwell

## Power and Sample Size Planning{.smaller}

```{r echo = F}
library(knitr)
library(papaja)
library(rio)
library(ggplot2)
library(dplyr)
library(broom)
library(kableExtra)
include_graphics("oprah.png")
```

## A Blender Mix{.smaller}

- Accuracy in Parameter Estimation and Simulation Approaches for Sample Size Planning
- How we took a bunch of interesting ideas and mixed them together

## Sample Size Planning{.smaller}

- Sample size planning is often thought of as "point and click" 
  - G\*Power: https://www.psychologie.hhu.de/arbeitsgruppen/allgemeine-psychologie-und-arbeitspsychologie/gpower
  - https://jakewestfall.shinyapps.io/pangea/ 
  - https://pwrss.shinyapps.io/index/
  - https://designingexperiments.com/
- Sample size planning is *technically* a closed-form solution for many analysis plans 
- An incredible number of cool *R* packages for sample size planning, such as *pwr*
- So, why do we need new innovations for power?

## The Need{.smaller}

- TOPS movement + pre-registration + grants + registered reports = need for power analyses 
- Power analyses are just our *best guesses* and are likely wrong 
- Many Analyst studies show us that there is no design = analysis answer
- The smallest effect of interest may be unknown
- Some research papers do not have one specific hypothesis (i.e., dataset creations)
- Once you leave the *t*-test behind, power becomes more complicated and often based on simulation

## Our Use Case{.smaller}

- Research studies that use many items to assess the parameter of interest 
- Research studies designed to collect data on many items and share the data 
- We should be careful not to assume all items are equal ... 
- And move away from using item-level averages as parameters of interest

## Combining Toolkits{.smaller}

- Accuracy in Parameter Estimation: finding the sample size that allows for "accurately measured" parameters 
  - Determine a "sufficiently narrow" confidence interval around your parameter
  - Determine the sample size that should provide that CI
- Bootstrapping (sort of) and Simulation
  - Taking pilot data and simulating various sample sizes based on bootstrapping your sample
  - Use this technique to find the sample size for a "sufficiently narrow" CI for items

## Sequential Testing{.smaller}

- Sequential testing: examine the parameter of interest for the intended CI
  - After each participant 
  - At regular intervals during data collection
- Benefits: 
  - Maximizing the usefulness of data collection 
- Cons: 
  - Usually requires code based skill sets

## Proposed Method{.smaller}

```{r table-summary, results = 'asis'}
kable(data.frame(
  "Step Number" = c("1", "2" , "3", "4", "5", "6"),
    "Proposed Steps" = c("Use representative pilot data.",
    "Calculate standard error of each of the items in the pilot data. Using the 40%, determine the cutoff and stopping rule for the standard error of the items.",
    "Create bootstrapped samples of your pilot data starting with at least 20 participants up to a maximum number of participants.", 
    "Calculate the standard error of each of the items in the bootstrapped data. From these scores, calculate the percent of items below the cutoff score from Step 2.",
    "Determine the sample size at which 80%, 85%, 90%, 95% of items are below the cutoff score. Use the correction formula to adjust your proposed sample size based on pilot data size, power, and percent variability.",
    "Report all values. Designate one as the minimum sample size, the cutoff score as the stopping rule for adaptive designs, and the maximum sample size.")
  ), 
  caption = "Proposed Procedure for Powering Studies with Multiple Items",
  col.names = c("Step", "Proposed Steps"),
  #align = c("m{1cm}", rep("m{6cm}", 2)),
  #font_size = "footnotesize"
) %>% 
  kable_styling(kable_input = ., 
                font_size = 25)
```

## Package{.smaller}

- Upcoming package `semanticprimeR` as part of a larger project 
- `devtools::install_github("SemanticPriming/semanticprimeR")`
- Functions for each step of the proposed process
- Functionality for when you have example data and when you do not (i.e., simulate example multiple-item data)
- As part of the manuscript and *semanticprimeR* package, we provide 12+ examples online 
- Psycholinguistics, social psychology, COVID related, traditional cognitive psychology

## Example: Step 1 (Pilot Sample){.smaller}

```{r import-data-example, include = F, echo = F}
elpDF <- import("../../data/ELPDecisionData.zip")
conDF <- import("../../data/concreteness_trial.rda")

# grab only the overlap for this fake study
elpDF <- elpDF %>% 
  filter(Stimulus %in% conDF$Word) %>% 
  filter(Type == 1)

conDF <- conDF %>% 
  filter(Word %in% elpDF$Stimulus)

# sample the data of with 20 words in each category
average_item_score <- conDF %>% 
  filter(Rating != "N" & Rating != "n") %>% 
  mutate(Rating = as.numeric(Rating)) %>% 
  group_by(Word) %>% 
  summarize(avgrating = mean(Rating))

con_words <- c(
  average_item_score %>% 
    filter(avgrating <= 2) %>% 
    sample_n(size = 20) %>% 
    pull(Word), 
  average_item_score %>% 
    filter(avgrating >= 4) %>% 
    sample_n(size = 20) %>% 
    pull(Word)
)

elp_small <- elpDF %>% filter(Stimulus %in% con_words)
# elp_small$Stimulus <- droplevels(elp_small$Stimulus) already a character

# figure out pilot size
pilot_size_e <- round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)))

# figure out data loss
elp_small$RT[elp_small$Accuracy == "0"] <- NA
data_loss_e <- elp_small %>% 
 group_by(Stimulus, Type) %>% 
 summarize(percent_correct = sum(!is.na(RT))/n(), .groups = "keep")
elp_use <- elp_small %>% filter(!is.na(RT))

# cutoff score 
SE_e <- tapply(elp_use$RT, elp_use$Stimulus, function (x) { sd(x)/sqrt(length(x))})
cutoff_e <- quantile(SE_e, probs = seq(from = .1, to = .9, by = .1))

elp_use <- elp_use %>% 
  ungroup() %>% 
  as.data.frame()
```

- You want to run a lexical decision project measuring response latencies for concrete and abstract words 
- You can use the English Lexicon Project as pilot data + previous publications of concreteness ratings 
- In these studies, we also have to factor in data loss!
  - Combined data includes `r length(unique(elpDF$Stimulus))` real words filtered down to `r length(unique(elp_small$Stimulus))` selected stimuli
  - Average sample size per word:  `r format(round(mean(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2), nsmall = 2)` (*SD* = `r round(sd(tapply(elp_small$RT, elp_small$Stimulus, length)), digits = 2)`)
  - Pilot sample size: *n* = `r pilot_size_e`

## Example: Step 2 (Calculate Cutoff){.smaller}

```{r echo = T}
library(semanticprimeR)
cutoff <- calculate_cutoff(population = elp_use, # pilot data or simulated data
  grouping_items = "Stimulus", # name of the item indicator column
  score = "RT", # name of the dependent variable column
  minimum = min(elp_use$RT), # minimum possible/found score
  maximum = max(elp_use$RT)) # maximum possible/found score
```

## Example: Step 2 (Calculate Cutoff){.smaller}

```{r echo = T}
cutoff$se_items # all standard errors of items
cutoff$sd_items # standard deviation of the standard errors
cutoff$cutoff # 40% decile score
cutoff$prop_var # proportion of possible variance 
```

## Example: Step 3 (BootSim Samples){.smaller}

```{r boot-sample, include = T, echo = T}
samples <- bootstrap_samples(start = 20, # starting sample size
  stop = 100, # stopping sample size
  increase = 5, # increase bootstrapped samples by this amount
  population = elp_use, # population or pilot data
  replace = TRUE, # bootstrap with replacement? 
  nsim = 500, # number of simulations to run
  grouping_items = "Stimulus") # item column label

head(samples[[1]])
```

```{r echo = F}
for (i in 1:length(samples)){
  samples[[i]] <- samples[[i]] %>% 
    ungroup() %>% 
    as.data.frame()
}
```

## Example: Step 4-5 (Calculate Proportion){.smaller}

```{r echo = F}
calculate_proportion <- function(samples,
                                 cutoff,
                                 grouping_items = NULL,
                                 score){

  if(is.null(samples)){ stop("You must include list of samples to examine.") }
  if(is.null(cutoff)){ stop("You must include the cutoff score for standard error.") }
  if(is.null(grouping_items)){ stop("You must include the grouping variable,
                                    which is normally the item number.") }
  if(is.null(score)){ stop("You must include the score or variable you are
                           estimating sample size for.") }

  summary_list <- list()
  score2 <- sym(score)

  # loop and calculate
  for (i in 1:length(samples)){

    summary_list[[i]] <- samples[[i]] %>%
      group_by(.data[[grouping_items]]) %>%
      summarize(se = sd(.data[[score]])/sqrt(n())) %>%
      ungroup() %>%
      summarize(percent_below = sum(se <= cutoff) / length(se),
                num_items = length(se)) %>%
      mutate(sample_size = nrow(samples[[i]]) / num_items)

  } # end loop and calculate

  # create end summary
  summary_DF <- bind_rows(summary_list) %>%
    select(-num_items) %>%
    group_by(sample_size) %>%
    summarize(percent_below = mean(percent_below))

  # return values
  return(summary_DF)

}

```

```{r calc-prop, include = T, echo = T}
proportion_summary <- calculate_proportion(samples = samples, # samples list
  cutoff = cutoff$cutoff, # cut off score 
  grouping_items = "Stimulus", # item column name
  score = "RT") # dependent variable column name

head(proportion_summary)
```

## Example: Step 6 (Apply Correction){.smaller}

```{r calc-correct, include = T, echo = T}
corrected_summary <- calculate_correction(
  proportion_summary = proportion_summary, # prop from above
  pilot_sample_size = pilot_size_e, # number of participants in the pilot data 
  proportion_variability = cutoff$prop_var, # proportion variance from cutoff scores
  power_levels = c(80, 85, 90, 95)) # what levels of power to calculate

corrected_summary
```

## Last Thoughts{.smaller}

- Use case: multiple items that intend on using item level focused analyses
- Should simulate only what is expected for a participant to do in the study 
  - Large numbers of items may bias estimates 
- Could combine with "traditional" power 
- Provides "well-measured" data --> not a specific decision for a specific sample

## Thanks{.smaller}

- Thanks for listening!
- Reproducible manuscript: https://github.com/SemanticPriming/stimuli-power 
- Package: https://github.com/SemanticPriming/semanticprimeR
- Scan me for a copy of this talk with links:

```{r include = T}
knitr::include_graphics("aps_24_qr.png")
```

## Simulation Method{.smaller}

- To evaluate our approach, we ran a simulation study: 
  - Scale size: popular cognitive scales (1-7 measurements, 0-100 percentage measurements, and 0-3000 response latency type scale data)
  - Item heterogeneity: small, medium, large
  - Skew: normal distributions versus skewed (ceiling) distributions 
  - Pilot sample size: 20 to 100 increasing in units of 10
- 1,620,000 simulations of 3 X 3 X 2 X 9 design

```{r table-sim}
#| fig-cap: " "
#| filters:
#| - parse-latex
#| 
kable(data.frame(
  Information = c("Minimum", "Maximum", "Mu", "Skewed Mu", "Sigma Mu", 
                  "Sigma", "Small Sigma Sigma", 
                  "Medium Sigma Sigma", "Large Sigma Sigma"),
  Likert = c(1, 7, 4, 6, .25, 2, .2, .4, .8),
  Percent = c(0, 100, 50, 85, 10, 25, 4, 8, 16),
  Milliseconds = c(0, 3000, 1000, 2500, 150, 400, 50, 100, 200)
), caption = "Parameter Values for Data Simulation",
escape = FALSE,
#format = "latex"
) %>% 
  kable_styling(kable_input = ., 
                font_size = 25)

```

## Simulation Results: Scale Size{.smaller}

```{r import-data}
summary_long <- bind_rows(
  import("../../simulation_study/simulation_middle/simulated_summary_data_middle.csv") %>% 
    mutate(data_type = "Normal"), 
    import("../../simulation_study/simulation_ceiling/simulated_summary_data_ceiling.csv") %>% 
    mutate(data_type = "Ceiling"))

summary_long$scale_size <- factor(summary_long$scale_size,
                                  levels = c("Likert", "Percent", "Milliseconds"))
summary_long$variability <- factor(summary_long$variability, 
                                   levels = c("Small Heterogeneity", 
                                              "Medium Heterogeneity", 
                                              "Large Heterogeneity"))

### make a decision here ### 
summary_long <- summary_long %>% filter(name != "Decile 0" & name != "Decile 10" & name != "Decile 20")
```

```{r scale-size-figure}

likert_var <- sqrt(((7-1)^2)/4)
percent_var <- sqrt(((100-0)^2)/4)
millisecond_var <- sqrt(((3000-0)^2)/4)

summary_long$percent_var <- ifelse(
  summary_long$scale_size == "Likert", summary_long$sd_item/likert_var, ifelse(
    summary_long$scale_size == "Percent", summary_long$sd_item/percent_var, 
      summary_long$sd_item/millisecond_var
  )
)

mean_var <- tapply(summary_long$percent_var, summary_long$scale_size, mean)

# New facet label names for power variable
power.labs <- c("80% of items", "85% of items",
                "90% of items", "95% of items")
names(power.labs) <- c("80", "85", "90", "95")

plot_colors <- c("#4FEAAC", "#FF826F", "#4F84FD")

ggplot(summary_long %>% 
         group_by(original_n, power, scale_size) %>% 
         summarize(sample_size_avg = mean(sample_size)),
       aes(original_n, sample_size_avg, color = scale_size)) + 
  geom_point() + 
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  geom_line(aes(original_n, original_n), color = "black") + 
  theme_linedraw(base_size = 15) + 
  xlab("\nPilot Sample Size") + 
  ylab("Projected Sample Size\n") + 
  scale_color_manual(name = "Scale Size",
                     values = c(plot_colors)) +
  facet_wrap(~power,
             labeller = labeller(power = power.labs))+
  theme_bw() +
  theme(legend.position="bottom")
```

## Simulation Results: Skew{.smaller}

```{r scale-skew-figure}

ggplot(summary_long %>% 
         group_by(original_n, power, data_type) %>% 
         summarize(sample_size_avg = mean(sample_size)),
       aes(original_n, sample_size_avg, color = data_type)) + 
  geom_point() + 
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  geom_line(aes(original_n, original_n), color = "black") + 
  theme_linedraw(base_size = 15) + 
  xlab("\nPilot Sample Size") + 
  ylab("Projected Sample Size\n") + 
  scale_color_manual(name = "Scale Skew",
                     values = c(plot_colors)) +
  facet_wrap(~power,
             labeller = labeller(power = power.labs))+
  theme_bw() +
  theme(legend.position="bottom")
```

## Simulation Results: Item Heterogeneity{.smaller}

```{r scale-hetero-figure}

ggplot(summary_long %>% 
         group_by(original_n, power, variability) %>% 
         summarize(sample_size_avg = mean(sample_size)),
       aes(original_n, sample_size_avg, color = variability)) + 
  geom_point() + 
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_line() + 
  geom_line(aes(original_n, original_n), color = "black") + 
  theme_linedraw() + 
  xlab("Pilot Sample Size") + 
  ylab("Projected Sample Size") + 
  scale_color_manual(name = "Item Heterogeneity",
                     values = c(plot_colors)) +
  facet_wrap(~power,
             labeller = labeller(power = power.labs))+
  theme_bw() +
  theme(legend.position="bottom")
```

## Dealing with Pilot Sample Size{.smaller}

- At some point, power usually asymptotes with increasing sample size 
- So, we need a correction:

$$ 1 - \sqrt{\frac{N_{Pilot} - min(N_{Simulation})}{N_{Pilot}}}^{log_2(N_{Pilot})}$$

## Dealing with Pilot Sample Size{.smaller}

```{r corrected-figure}

decay <- 1-sqrt((summary_long$original_n-20)/summary_long$original_n)^log2(summary_long$original_n)
summary_long$new_sample <- summary_long$sample_size*decay
ggplot(summary_long %>% 
         group_by(original_n, power, variability) %>% 
         summarize(sample_size_avg = mean(sample_size),
                   new_sample_size_avg = mean(new_sample)),
       aes(original_n, new_sample_size_avg, color = variability)) + 
  geom_line(aes(original_n, original_n), color = "black") +
  geom_point(aes(original_n, original_n), color = "black") + 
  geom_point() + 
  geom_line() + 
  xlab("\nPilot Sample Size") + 
  ylab("Projected Sample Size\n") + 
  scale_color_manual(name = "Item Heterogeneity",
                     values = c(plot_colors)) +
  facet_wrap(~power,
             labeller = labeller(power = power.labs))+
  theme_bw() +
  theme(legend.position="bottom")
```

## Researchers Have *One* Sample{.smaller}

- Long story short: we can provide a function for researchers to use to control pilot sample size
- We also determined *which* level "sufficiently small" was probably best

```{r R2-cutoff, include = F, echo = F}
by_cutoff <- list()
R2 <- list()
for (cutoff in unique(summary_long$name)){
 by_cutoff[[cutoff]] <- lm(new_sample ~ sample_size + percent_var + original_n + power + log2(sample_size) + log2(percent_var) + log2(original_n) + log2(power), data = summary_long %>% filter(name == cutoff))
 R2[cutoff] <- summary(by_cutoff[[cutoff]])$r.squared
}
```

```{r table-decile}
table_correct <- tidy(summary(by_cutoff$`Decile 40`))
table_correct <- table_correct[c(1, 2, 4, 6, 8, 9, 3, 7, 5) , ]
table_correct$term <- c("Intercept", "Projected Sample Size", "Pilot Sample Size", 
                        "Log2 Projected Sample Size", "Log2 Pilot Sample Size", 
                        "Log2 Power", "Proportion Variability", 
                        "Log2 Proportion Variability", "Power")
table_correct$estimate <- printnum(table_correct$estimate, digits = 3)
table_correct$std.error <- printnum(table_correct$std.error, digits = 3)
table_correct$statistic <- printnum(table_correct$statistic, digits = 3)
table_correct$p.value <- printnum(table_correct$p.value, digits = 3,
                 zero = FALSE, gt1 = F)
kable(table_correct, 
     col.names = c("Term", "Estimate", "SE", "t", "p"),
     caption = "Parameters for 40% Decile Cutoff Scores") %>% 
  kable_styling(font_size = 25)
```