### 2024-11-11 simulation_study/simulate_summary_bimodal.Rmd

---
title: "Methods Power Study - Bimodal Simulations"
author: "Erin M. Buchanan"
date: "Last Knitted: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r message = F}
set.seed(895893) # 233232
library(dplyr)
library(ggplot2)
library(tidyr)
library(semanticprimeR)
```

This document examines a test of what a researcher might find if they had bimodal distributions on their data. We will simulate a pilot study by generating normal, floor, and ceiling scores. From those populations, we will select a mix of items that are "normal" and a mix of items that are bimodal (by joining together the ceiling and floor effects with a .5 probability). Finally, we will calculate the rest of the steps proposed as normal.

```{r simulate-bimodal}
# simulate population 
normal_pops <-semanticprimeR::simulate_population(mu = 4, # item means
  mu_sigma = .2, # variability in item means 
  sigma = 2, # item standard deviations
  sigma_sigma = .2, # standard deviation of the standard deviations
  number_items = 30, # number of items
  number_scores = 1000, # number of participants
  smallest_sigma = .02, #* smallest possible standard deviation
  min_score = 1, #* minimum score for truncating purposes
  max_score = 7, #* maximum score for truncating purposes
  digits = 0) #* number of digits for rounding

floor_pops <- semanticprimeR::simulate_population(mu = 2, # item means
  mu_sigma = .2, # variability in item means 
  sigma = 1, # item standard deviations
  sigma_sigma = .1, # standard deviation of the standard deviations
  number_items = 30, # number of items
  number_scores = 1000, # number of participants
  smallest_sigma = .02, #* smallest possible standard deviation
  min_score = 1, #* minimum score for truncating purposes
  max_score = 7, #* maximum score for truncating purposes
  digits = 0) #* number of digits for rounding

ceiling_pops <- semanticprimeR::simulate_population(mu = 6, # item means
  mu_sigma = .2, # variability in item means 
  sigma = 1, # item standard deviations
  sigma_sigma = .1, # standard deviation of the standard deviations
  number_items = 30, # number of items
  number_scores = 1000, # number of participants
  smallest_sigma = .02, #* smallest possible standard deviation
  min_score = 1, #* minimum score for truncating purposes
  max_score = 7, #* maximum score for truncating purposes
  digits = 0) #* number of digits for rounding

# simulate sample 
item_number <- 1:30
prop_mix <- .5 # manipulated this number 
pilot_sample_size <- 30
bimodal_items <- sample(item_number, size = round(length(item_number)*prop_mix), replace = FALSE)

researcher_sample <- 
  bind_rows(
    floor_pops %>% 
      filter(item %in% bimodal_items) %>% 
      group_by(item) %>% 
      slice_sample(n = pilot_sample_size / 2), 
    ceiling_pops %>% 
      filter(item %in% bimodal_items) %>% 
      group_by(item) %>% 
      slice_sample(n = pilot_sample_size / 2),
    normal_pops %>% 
      filter(!(item %in% bimodal_items)) %>% 
      group_by(item) %>% 
      slice_sample(n = pilot_sample_size)
  )

library(ggridges)

ggplot(researcher_sample %>% 
         filter(item < 11), aes(x = score, y = factor(item), fill = factor(item))) +
  geom_density_ridges(alpha = 0.7, scale = 1) +
  labs(x = "Score", y = "Item") +
  theme_minimal() +
  theme(legend.position = "none") + 
  coord_cartesian(xlim = c(1,7))

# researcher_sample %>% 
#   group_by(item) %>% 
#   summarize(n = n(),
#             mean = mean(score),
#             sd = sd(score), 
#             min = min(score),
#             max = max(score))
# 
# researcher_sample %>% 
#   filter(item %in% bimodal_items) %>% 
#   group_by(item) %>%
#   table()

# calculate cutoff 
cutoff <- calculate_cutoff(population = researcher_sample, # pilot data or simulated data
  grouping_items = "item", # name of the item indicator column
  score = "score", # name of the dependent variable column
  minimum = 1, # minimum possible/found score
  maximum = 7) # maximum possible/found score

# bootstrap samples
samples <- bootstrap_samples(start = 20, # starting sample size
  stop = 100, # stopping sample size
  increase = 5, # increase bootstrapped samples by this amount
  population = researcher_sample, # population or pilot data
  replace = TRUE, # bootstrap with replacement? 
  nsim = 1000, # number of simulations to run
  grouping_items = "item") # item column label

# calculate proportions
proportion_summary <- calculate_proportion(samples = samples, # samples list
  cutoff = cutoff$cutoff, # cut off score 
  grouping_items = "item", # item column name
  score = "score") # dependent variable column name

# calculate correction
corrected_summary <- calculate_correction(
  proportion_summary = proportion_summary, # prop from above
  pilot_sample_size = 30, # number of participants in the pilot data 
  proportion_variability = cutoff$prop_var, # proportion variance from cutoff scores
  power_levels = c(80, 85, 90, 95)) # what levels of power to calculate

rio::export(corrected_summary, paste0("simulation_bimodal/summary_", 
                                 prop_mix, ".csv"), 
                                 row.names = F)
```

```{r}
list_files <- list.files("simulation_bimodal/",
                         pattern = "*.csv",
                         full.names = TRUE)

import_files <- lapply(list_files, rio::import)

full_data <- bind_rows(import_files) %>% 
  mutate(size = rep(c(.1, .2, .3, .4, .5, .6, .7, .8, .9, 0, 1), each = 4))

rio::export(full_data, "simulation_bimodal/summary_bimodal.csv", row.names = F)
```